\documentclass[a4paper,12pt]{article}
\usepackage[margin=.7in]{geometry}

% Encoding and font settings
\usepackage[utf8]{inputenc}

% Math packages
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{bbm}
% Formatting packages
\usepackage{indentfirst,color,fullpage}

% Graphics and tables
\usepackage{graphicx}
\usepackage[flushleft]{threeparttable}
\usepackage{booktabs}

% Hyperlinking and referencing
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{url}

% Additional functionality
\usepackage{footnote}
\usepackage{xcolor}
\usepackage{bm}
\usepackage{ulem}
\newcommand{\red}{\textcolor{red}}
\newcommand{\purple}{\textcolor{purple}}

\begin{document}

\noindent \textbf{{\Large Response to Reviewers:\\
``MOPED: A moving sum method for change point detection in pairwise extremal dependence''}}

\vspace*{.25cm}

\noindent \textbf{{\large(Manuscript ID: Extremes b660ab30-cacb-4a14-a6a9-e60fcf48c005)}}

\vspace*{.25cm}



\noindent {\it by Euan T.~McGonigle, Matthew Pawley, Jordan Richards\footnote{Corresponding author: Jordan Richards \\
\hspace*{.75cm} School of Mathematics, University of Edinburgh, UK\\
\hspace*{.75cm} \href{mailto:jordan.richards@ed.ac.uk}{jordan.richards@ed.ac.uk}} and Christian Rohrbeck}

\vspace{.5cm}

\noindent {\it To: Rafal Kulik, Ph.D.} \\
{\it Editor, Extremes} \\
{\it Chen Zhou, Ph.D.} \\
{\it Guest Editor
} \\

Dear Rafal and Chen, \\

Thank you for the opportunity to provide an updated version of our manuscript entitled ``MOPED: A moving sum method for change point detection in pairwise extremal dependence'' for consideration in Extremes. The comments from the reviewers have been very helpful, and we have addressed them thoroughly.\\

Major changes to the text have been highlighted in \textcolor{red}{red}.
\clearpage
\newpage

\clearpage
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%% End of Highlights %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Editor %%% Editor %%% Editor %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent {\Large \bf {Responses to the Guest Editor}}\\

\noindent \textcolor{violet}{I have received reports from two experts in the field. Referee 1 recommended "Revise" with a list of comments while Referee 2 is more critical.\\
The main critique of Referee 2  is that the paper is lack of theoretical guarantee. To my view, it is also important to have methodological papers ahead of theoretical work. Therefore, I would not emphasize on developing new theories. Instead, it is sufficient to provide arguments why the proposed procedure has the potential to achieve theoretical guarantee.\\
With this idea in mind, I'd like to invite the authors for a revision addressing the comments of Referee 1 and the minor comment of Referee 2.}\\

\noindent \textcolor{violet}{\bf Response}: Thank you for these comments. We have thoroughly addressed the comments of Reviewer 1, and so the latest version of the manuscript includes two additional simulation studies, a re-write of the methodology section, and additional comments on model assumptions and justification throughout the paper. Per your advice, we have opted to leave the major comments of Reviewer 2 unaddressed, but we have addressed all of the minor comments. \textbf{Note that we have not derived any new theoretical results in this updated manuscript.}\\


\clearpage
\newpage



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Reviewer 1 %%% Reviewer 1 %%% Reviewer 1 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent {\Large {\bf Responses to Reviewer 1}}\\

\noindent \textcolor{teal}{The paper is well-written, interesting, and potentially useful. However, several assumptions are not stated explicitly, some imported results are used
without verifying their conditions in the present setting, and the theoretical
guarantees (e.g., consistency for break detection) are not yet articulated.
Clarifying these points and expanding the Monte Carlo section would substantially
strengthen the paper.}



%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%
\section*{{Major Comments}}



%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{1. The way you introduce the multivariate time series (viz., Eq. (4)) reads as if a break affects the entire joint distribution, not just extremal
dependence. Please clarify.}\\

\noindent \textcolor{teal}{\bf Response}: See below. \\


%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{2. The authors impose several assumptions within their text. I would
prefer if assumptions were made more explicit using e.g.
\begin{itemize}
\item \textbf{Assumption 1} [Margins]
\item \textbf{Assumption 2} [(No) Serial dependence]
\item \textbf{Assumption 3} [Moments]
\end{itemize}
Also, state the null hypothesis and alternative explicitly.}\\

\noindent \textcolor{teal}{\bf Response}:  Thank you for these comments. We have added the following text below Eq. (5)\\

\textcolor{red}{To further clarify our proposed change point model, we reiterate our assumptions on the time series $\{\mathbf{X}_t\}^n_{t=1}$:
\begin{itemize}
\item \textbf{Assumption 1:}  Marginally, ${X}_{i,t} \in \mbox{RV}^1_+(2)$ for all $i=1,\dots,d,t=1,\dots,n.$ [Margins]
\item \textbf{Assumption 2:}  $\mathbf{X}_t \in \mbox{RV}^d_+(2),t=1,\dots,n,$ are independent. [Serial independence]
\item \textbf{Assumption 3:}  There exist a collection of $q+2$ (with $q \geq 0$) change points  $\mathcal C = \{\tau_l \}_{l=0}^{q+1},$ with $\tau_0=0$ and $\tau_{q+1}=n$, such that the TPDM of the regularly varying $\mathbf{X}_t$ is $\Sigma^{(l)}$ for $ \tau_{l-1} < t \leq \tau_{l},l=1,\dots,q,$ with $\Sigma^{(l)}\neq \Sigma^{(l')}$ for all $l\neq l'$. [Change points]
\end{itemize}}

\textcolor{red}{\textit{Remark}: As an implication of \textbf{Assumption 1}, the $\beta$-th moment of $X_{i,t}$ is finite only for $\beta< 2$. Note that $\mathbf{X}_t \in \mbox{RV}^d_+(2)$ is an asymptotic statement, and we make no assumptions on the sub-asymptotic joint distribution of $\mathbf{X}_t \in \mbox{RV}^d_+(2)$. We assume that only the corresponding TPDM, and thus the pairwise extremal dependence, differs across change points. }\\
Note that our focus is on the problem of detecting and estimating multiple change points, rather than the closely related testing problem, so that we do not have a null or alternative hypothesis statement. The number of change points, $q\geq 0$, is an estimable parameter in our model, and thus we have framed the hypothesis test as an estimation problem.  \\

%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%

\noindent \textcolor{teal}{3. In several places you rely on results from the literature. It is not clear,
at least to me, whether this is always justified here. For example,
on page 7, the proposed procedure borrows techniques from Eichinger
and Kirch (2018, Bernoulli), who impose specific regularity conditions.
Please state explicitly which results you use and state the corresponding
conditions; verify that these conditions hold in your current setup.}\\

\noindent \textcolor{teal}{\bf Response}: Thank you for raising this important point. To clarify: we adopt the MOSUM detection methodology from \cite{eichinger2018}, but we do not rely on their theoretical results or consistency guarantees.

Our contribution is empirical and methodological in nature. We do not make theoretical claims about detection consistency or invoke the regularity conditions required for the asymptotic results in \cite{eichinger2018}. Rather, we demonstrate the practical utility of adapting the MOSUM approach to the context of extremal dependence changes through numerical studies and the EEG application.

In the Discussion section, we note that rigorous theoretical justification for change point detection in extremes remains an open area for future research.  \\

%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{4. More generally, some theoretical evidence in terms of a formal result
(e.g. in form of a theorem) confirming that the proposed procedure is
able to, for example, consistently detect breakpoints would be welcome.}\\

\noindent \textcolor{teal}{\bf Response}: We agree that formal theoretical guarantees would strengthen the contribution. However, establishing consistency results for change point detection in extremal dependence is beyond the scope of the current work and represents a significant open theoretical challenge.

Standard change point detection techniques, such as those in \cite{eichinger2018}, rely on careful decomposition of test statistics in the presence of change points. Extending these techniques to our setting would require new technical results involving explicit convergence rates in extreme value theory which, to the best of our knowledge, are not currently available in the literature. We view the development of rigorous theoretical foundations for change point detection in extremes as an important direction for future research. \\

%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{5. Given that you work with time series, the assumption of independence
feels restrictive. Can it be relaxed to allow for some amount of dependence?
For example, you are citing Hoga (2018, Biometrika), whose method is applicable to short memory ($\beta$-mixing) processes. Alternatively,
you might discuss a prior filtering step; i.e., your procedure
would be applied to residuals from estimated marginal time series models.
Here, the cited paper of B\"ucher, J\"aschke and Wied (2015, JoE)
contains a preliminary discussion; in a related paper, Borsch, Mayer
and Wied ('Consistent estimation of multiple breaks \dots ', 2024, JBES)
provide also some theoretical results in that respect for multiple break
point detection for copulas. Of course, a technical treatment seems
beyond the scope, but at least a discussion seems warranted. Adding
a short simulation comparing e.g. IID and AR-GARCH settings could
also be very informative here. This is the more important because in your application (p. 16) it seems as if you apply some (pre-)filtering..}\\

\noindent \textcolor{teal}{\bf Response}: This is a great point. We have added an additional simulation study, Scenario 3, which considers change point detection for a vector autoregressive model. We have added the below details to Section 3.2:
 \textcolor{red}{For Scenario 3, we generate data from a first-order vector autoregressive model:
 $$\mathbf{X}_t=\Phi \mathbf{X}_{t-1}+ \boldsymbol{\epsilon}_t,\quad t = 1,\dots,n,$$
 where $\Phi = \phi I_{d\times d}$ for autoregressive (AR) parameter $\phi \in[0,1)$ and where $\boldsymbol{\epsilon}_t$ is $d$-variate Gaussian with zero mean, unit variance, and correlation matrix $\Omega$. The correlation matrix $\Omega$ of $\boldsymbol{\epsilon}_t$ differs across the change points, and it is parametrised as described above for Scenario~1. }

The results are stored in Tables A9--A14 of the Appendix, and summarised at the end of Section 3.2:
\textcolor{red}{For Scenario 3, Tables~A9--A11 give the results for $d=2$ while Tables~A2--A14 give the results for $d=8$. As the value of the autocorrelation parameter $\psi$ increases from $0$ to $0.6$, we observe the expected degradation in the performance of all considered change point detection methods: MOPED, E-divisive, and the parametric approach of \cite{hazra2025estimating}. For all methods, we observe a reduction in estimation accuracy for $q$, the number of change points, as well as the average covering metric and V-measure. However, we note that the extremal change point methods - MOPED and \cite{hazra2025estimating} - exhibit relative robustness to the presence of autocorrelation in comparison with E-divisive. When $\psi=0$, i.e., the data are independent, E-divisive outperforms MOPED in some cases (Tables~A9 and A12). However, even under mild autocorrelation (Tables~A10 and A13), the performance of E-divisive is markedly worse than MOPED, with E-divisive routinely overestimating the number of change points by at least two. Conversely, MOPED facilitates reasonable estimation accuracy even under strong autocorrelation (Tables~A1 and A4), albeit with some evidence that the true number of change points is underestimated.}

We also note in the application:
\textcolor{red}{To remove serial dependence in these data, we sub-sample every 256th value, where 256Hz is the sampling rate, and thus we sub-sample every second. While this does not completely remove the autocorrelation in the data, the results from Scenario 3 of the simulation study (Section~3) suggest that MOPED is relatively robust to the presence of mild autocorrelation.}\\
and comment in Section 2.5:\\
\textcolor{red}{\textit{Remark}: if independence of observations cannot be assumed, then alternative resampling methods that are applicable to the times series setting could be employed instead; for example the block bootstrapping approach of \cite{ng2022bootstrap}, or the wild bootstrap of \cite{shao2010dependent}.}\\


%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{6. At the very beginning you mention that margins are assumed to be unaffected
from breaks. First, although it seems to be a common assumption
in the literature, a discussion of when this is reasonable and how
violations would affect inference would be appreciated (see e.g. Section
3.7 in B\"ucher, J\"aschke and Wied (2015, JoE)). Second, comment on
whether the assumption is plausible in your empirical application.}\\

\noindent \textcolor{teal}{\bf Response}: Thank you for this point. MOPED assumes that the margins are regularly varying with index $\alpha=2$, which is a less strict assumption than the margins being completely unaffected by the change point - rather we just require that the marginal regular variation condition is satisfied. 

For our application, previous studies have investigated the temporal non-stationarity in the marginal extremal behaviour, and found that this does exhibit a change point. As we have not explicitly accounted for this in our manuscript, we cannot guarantee that our MOPED estimates are due to changes in only the extremal dependence structure. We have commented on this limitation in the main text:\\
\textcolor{red}{As discussed by \cite{Guerrero2023}, these data exhibit temporal non-stationarity in their marginal extremal behaviour. For simplicity, we have not explicitly modelled the data margins. As such, the following estimated change points may reflect changes in the marginal extremal behaviour and/or the extremal dependence structure. However, MOPED still provides a quick and convenient exploratory tool for identifying structural changes in the extremal behaviour of these data and, as we are to show, also provides estimates that are in better agreement, relative to the competitor (E-divisive), with expert opinion. }


%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{7. On the Monte Carlo experiment:}\\

\noindent \textcolor{teal}{7a. To get a better feeling for the performance of the procedure, also smaller sample sizes than the one chosen (n = 5,000) should be
considered.}\\


\noindent \textcolor{teal}{\bf Response}: Thank you for this point. For brevity, we have only repeated the experiments for Scenario 1 - these are not given in Tables A3--A5 of the Appendix. We have added the accompanying text in Section 3.2:

 \textcolor{red}{For Scenario 1, we repeat the numerical experiment with a lower sample size of $n=2000$. In this setting, to reflect the lower sample size, we use a smaller bandwidth of $G=500$ for MOPED, and sequence of bandwidths $\mathcal{H}=\{300,400,500\}$ for MMMOPED. The results are reported in Tables~A3--A5. We observe generally poorer performance of all methods (relative to $n=5000$), but MOPED retains its competitiveness with E-divisive and the parametric detection method \citep{hazra2025estimating}. We also observe that MOPED exhibits robustness to the sample size in cases where change points are easier to detect, i.e., $\rho = 0.6$. Note that, in our real data application (Section~4), the lowest sample size that we consider is $n\approx 3500$.}\\


\noindent \textcolor{teal}{7b. Maybe I am missing something, but the performance of all methods in Scenario 2 is rather poor. Correct detections (column $\hat{q}-q$)
are on average the exception. Can you explain? Is the break-size (too) low relative to window/sample size?.}\\


\noindent \textcolor{teal}{\bf Response}: Thank you for this comment. In this case, the change in dependence structure is very subtle - the sub-asymptotic dependence structure, e.g., correlation, does not change, but the limiting extremal dependence structure will transition from asymptotic dependence to asymptotic independence. Despite of this subtlety, MOPED and MMMOPED perform well when the dimension is relatively large, i.e., $d>2$. However, all methods perform poorly when $d=2$. We have added the following comments to the simulation study.

\textcolor{red}{In contrast with Scenario 1, the change in the distribution across the change point is more subtle in Scenario 2; the data generating copula changes from one which exhibits asymptotic dependence to one which exhibits asymptotic independence, but with the (sub-asymptotic) correlation structure remaining constant across the change point. In the case where the change point is particularly difficult to identify, for example, when the dimension is low ($d=2$), all methods underestimate the true number of change points $q$.  However, Tables~A6 and A7 illustrate that the fixed-parameter MOPED method generally outperforms both E-divisive and the method of \cite{hazra2025estimating}.}\\

\noindent \textcolor{teal}{7c. On page 11 you state that you chose two significance levels $\alpha \in \{0.05, 0.10\}$ for the permutation testing. I cannot see which of
them is used later in the simulations. Some break point procedures
require $\alpha:=\alpha(n)\rightarrow0$ ad $n\rightarrow \infty$ for consistently estimating
the number of breaks (e.g. Proposition 8 in Bai and Perron 'Estimating
and testing linear models \dots', Econometrica, 1998). Could
you comment on why this is here not needed?}\\


\noindent \textcolor{teal}{\bf Response}: Thank you for catching this: we use the significance level $\alpha = 0.1$ in the simulations, which has now been clarified in the main text in Section 3.1. 

The permutation test (if completely enumerated for all $n!$ permutations) is exact for a finite sample, and controls the proportion of false positives (any change point declared when there are none) at the prescribed level $\alpha$. We demonstrate in the numerical experiments that the permutation test is able to achieve close to the nominal size level when there are no change points present for a variety of sample sizes.

%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%
\section*{{Minor Comments}}



%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{1. On page 7 you argue that ‘asymptotically, the quantity $\sigma_{i,j}$ for $i=j$ is one’. Please clarify what the limit is taken over. For example, $\sigma_{i,j}$ seems not to depend on the sample size $n$.}\\

\noindent \textcolor{teal}{\bf Response}: Thank you for pointing this out. We have replaced ``asymptotically'' with ``theoretically''.\\ 


%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{2. The introductory Section 2.1. to the method is technically quite dense. I would welcome if the mathematically concepts could be presented in
a more accessible way.}\\

\noindent \textcolor{teal}{\bf Response}: Thank you for this point. We have re-written the opening of Section 2.1, which is provided below.  \\

  \textcolor{red}{A $d$-dimensional random vector $\mathbf{X}\in \mathbbm{R}^d_+$ is regularly varying with tail index $\alpha>0$, denoted by $\mathbf{X}\in {\rm RV}^d_+(\alpha)$, if there exists a sequence $b_n\to \infty$ such that $n\Pr(b_n^{-1}\mathbf{X} \in\cdot)\xrightarrow{v} v(\cdot)$ as $n\to\infty$, where the limiting Radon measure $v(\cdot)$ is supported on the cone $\mathbbm{E}^d_+:=[0,\infty]^d\setminus{\{\mathbf{0}\}}$ and where $\xrightarrow{v}$ denotes vague convergence; see, e.g., \cite{resnick2007heavy}. As a result of this vague convergence statement, the limit measure $v(\cdot)$ satisfies $(-\alpha)$-homogeneity, i.e., $v(r\mathcal{B})=r^{-\alpha}v(\mathcal{B})$ for any $r>0$ and any Borel subset $\mathcal{B}\subset \mathbbm{E}^d_+$. Courtesy of the $(-\alpha)$-homogeneity of  $v(\cdot)$, we can make $v(\cdot)$ amenable to modelling via a radial-angular decomposition of the limit measure into a product measure, defined on $\mathbb{R}_+ \times \mathcal{S}^{d-1}_+$, where {$\mathcal{S}^{d-1}_+:=\{\mathbf{x}\in \mathbbm{R}_+^d: ||\mathbf{x}||_2=1\}$} is the positive part of the unit $(d-1)$-sphere: for $r>0$, we consider $${v(\{\mathbf{x}\in \mathbbm{E}^d_+: ||\mathbf{x}||_2\geq r, \mathbf{x}/||\mathbf{x}||_2\in \mathcal{B}_H \})}=r^{-\alpha}H(\mathcal{B}_H),$$ where $\mathcal{B}_H \subset\mathcal{S}^{d-1}_+ $ is a Borel subset and where $H(\cdot)$ is an angular mass measure supported on $\mathcal{S}^{d-1}_+$. Loosely speaking, the angular measure $H(\cdot)$ determines the direction of a multivariate extreme event, while the measure $r^{-\alpha}$ determines the relative magnitude.} \\
  


%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%
\noindent \textcolor{teal}{3. The authors opt for a MOSUM detector. Personally, I would have
appreciated a discussion of why the authors did not choose a CUSUM
statistic using e.g. the procedure of Vostrikova (1981) (see also, more
recently, the wild binary segmentation of Fryzlewicz (‘\textit{Wild binary} \dots ’,
2014, Annals of Statistics)).}\\

\noindent \textcolor{teal}{\bf Response}: Thank you for this helpful suggestion. Change point methods based on applying test statistics in a localised manner, such as those based on MOSUMs or CUSUMs, have benefits and drawbacks, and we would not advocate that there is a uniformly best approach for change point detection. We selected the MOSUM approach for several reasons that are particularly relevant to our application. 

One benefit of the MOSUM approach is that it provides a straightforward method for simultaneous estimation of multiple change points using the local maxima of the test statistics, without the need for careful choice of search intervals as required in CUSUM-based approaches.

A key advantage of the MOSUM approach over cumulative sum (CUSUM) approaches is the use of symmetric windows, which is particularly important for a test statistic based on threshold exceedances since it ensures that the same number of threshold exceedances occur in the left and right windows, which is not the case for CUSUM approaches. This ensures the "effective sample size" is equal in both windows.

Lastly, the MOSUM approach is computationally more efficient than CUSUM-based approaches in our implementation; for a given bandwidth the permutation test only needs to be performed once regardless of the number of change points, whereas in CUSUM approaches, it would need to be performed on each search interval.

We have added a summary of this discussion at the end of Section 2.4: \\

\textcolor{red}{We remark that, whilst here we use the MOSUM-based approach,  alternatives based on cumulative sums (CUSUMs) such as the wild binary segmentation algorithm \citep{piotr2014} are also possible. The MOSUM approach, however, offers several advantages relevant to our application: the use of two symmetric windows of size $G$ in the computation of the test statistic allows for a natural use of order statistics to set the exceedance threshold (see discussion on extremal threshold selection in Section 3.1). This ensures that the same number of exceedances contribute in both windows, which helps with the empirical behaviour of the test statistic. Secondly, the MOSUM approach offers superior computational efficiency when combined with resampling methods (see Section 2.6), since the resampling procedure only need be performed once for a given bandwidth $G$.} \\

We have also added a brief motivation for MOSUM's usage in change point detection at the start of Section 2.4.\\ 

\textcolor{red}{The moving sum (MOSUM) methodology has successfully been applied to a variety of change point testing \citep{chu1995mosum} and estimation problems \citep{eichinger2018}}.\\

\noindent \textcolor{teal}{4. On page 6, there is an a to much, i.e. ‘of a a candidate’.}\\

\noindent \textcolor{teal}{\bf Response}: Fixed.\\ 

\clearpage
\newpage



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Reviewer 2 %%% Reviewer 2 %%% Reviewer 2 %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent {\Large {\bf Responses to Reviewer 2}}\\

\noindent \textcolor{blue}{The manuscript proposes a nonparametric method, named MOPED, for detecting multiple
change points in pairwise extremal dependence in multivariate time series. Pairwise
extremal dependence is summarised through the Tail Pairwise Dependence Matrix
(TPDM), see Cooley and Thibault (2019), which acts as a covariance-type summary for
multivariate heavy-tailed distributions.
For a moving window of width G, the authors compute left and right local estimates of
the TPDM and measure their discrepancy using a Moving Sum type statistic. Significant
peaks above a data-driven threshold, obtained via a permutation procedure under the
null of no change and independence, yield potential change points. The authors further
propose a multiscale and multi-threshold extension combining multiple bandwidths and
multiple extremal thresholds.
A simulation study compares the method with established procedures (E-divisive and,
in the bivariate case, the parametric method of Hazra et al.). Finally, an application to
multichannel neonatal EEG data demonstrates that the procedure can detect structural
changes associated with seizure episodes.}\\


%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%

\section*{{Strengths}}

%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%

\noindent \textcolor{blue}{3.1. Typos and language:
\begin{itemize}
\item The topic is highly relevant to Extremes: detecting changes in multivariate tail
dependence is an important and timely problem.
\item The idea of applying a MOSUM detector to local TPDM estimates is natural,
interpretable, and well motivated.
\item The method is nonparametric and applicable in moderately high dimensions.
\item The empirical work is extensive and generally well executed, including a diverse set
of simulation scenarios.
\item The EEG application is convincing and demonstrates the practical usefulness of the
approach.
\item Availability of an R package enhances reproducibility and practical impact.
\end{itemize}
}

\noindent \textcolor{blue}{\bf Response}: Thank you for these positive comments.\\


%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%

\section*{{Weaknesses and Limitations}}

\subsection*{{Major concern: lack of theoretical guarantees}}

%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%

\noindent \textcolor{blue}{The most significant limitation of the manuscript is the absence of fundamental theoretical
results regarding the proposed procedure. In particular:
\begin{itemize}
\item No consistency result is provided for the estimation of the number or locations of
change points.
\item No rate of convergence is established for the estimated change-point positions.
\item The asymptotic behaviour of the MOSUM-TPDM detector is not analysed, even
under simplified conditions.
\item The validity of the permutation test, crucial to the procedure, is not justified in the
context of multivariate extremes.
\item The method relies on an independence assumption that is unrealistic for many
applications.
\end{itemize}
For a methodological article submitted to Extremes, it is essential to provide at least
a partial theoretical framework. In its present form, the contribution remains essentially
empirical.
}

\noindent \textcolor{blue}{\bf Response}: Per the Editors comments, we have opted not to introduce any new theoretical results to our work, which we believe stands alone as a useful computational framework for identifying structural changes in the extremal dependence of time series. While theoretical support for MOPED is desirable, this is outside of the scope of the paper.\\

\subsection*{{Additional limitations}}

%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%

\noindent \textcolor{blue}{
\begin{itemize}
\item  The procedure is sensitive to tuning parameters $(G,k)$, despite the multiscale extension.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Thank you for this point. Sensitivity of the MOPED estimator to the choice of hyperparameters is, of course, a limitation of our method, which was our motivation for introducing the multiscale MOPED procedure which partially addresses this issue by allowing for adaptivity to multiscale change point scenarios.}
\item The focus on pairwise dependence means that changes in higher-order extremal
structure may remain undetected. \\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Added to Discussion: In this context, it is important to note that our approach is not able to detect all types of \textcolor{red}{changes in extremal dependence, due to the TPDM only considering pairwise tail dependence. Further work may develop change point algorithms designed to detect higher-order changes in extremal dependence by, for example, identifying changes in the angular measure.} } 
\item Minor issues in notation and presentation (e.g., TPDM/TDPM inconsistencies,
typographical errors). \\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{The TPDM/TDPM inconsistency has been fixed and minor typos have been addressed throughout the manuscript.}
\end{itemize}
}

%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%
%%% Section %%% Section %%% Section %%%

\section*{{Minor Corrections and Suggestions}}



%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%

\noindent \textcolor{blue}{3.1. Typos and language:
\begin{itemize}
\item “timesseries” should be corrected to “time series”.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Fixed.}\\
\item Ensure consistent spelling: “nonstationary” vs. “non-stationary”.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Fixed.}\\
\item In the abstract and introduction, harmonise the use of capitalisation for “MOving
sum” or use “moving-sum-based”.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Fixed.}\\
\item Replace any occurrences of “TDPM” by the correct acronym “TPDM”. \\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Fixed.}
\item Minor style improvements: use “it is advisable” instead of “it is advised”.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Fixed.}
\end{itemize}
}


%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%

\noindent \textcolor{blue}{3.2. Notation and clarity:}
\begin{itemize}
\item Add a clarification that margins are transformed empirically to Pareto(2) and that
the TPDM is (asymptotically) invariant under marginal transformations.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Added. \textcolor{red}{Note that $\mathbf{X}=(X_1,\dots,X_d) \in \mbox{RV}^d_+(\alpha)\Rightarrow X_i \in \mbox{RV}^1_+(\alpha), i=1,\dots,d,$ and that the TPDM is invariant under marginal transformations of $\mathbf{X}$ which conserve $\mbox{RV}^1_+(\alpha)$. }}
\item Explicitly state that the diagonal of the TPDM equals 1 asymptotically.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Added.}
\item In Section 2.3, clarify whether changes occur in the full angular measure or only in
pairwise dependence.\\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{Added.}

\item In the multiscale algorithms, explicitly note that the indexing proceeds from smallest
to largest k (or G).\\
\noindent \textcolor{blue}{\bf Response}: Algorithms 2 and 3 now have a note saying that we consider the change point values in increasing order with respect to $k$ or $\mathcal{G}$, respectively.
\end{itemize}

%%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%% Item %%%

\noindent \textcolor{blue}{3.3. Figures and tables:
\begin{itemize}
\item Add bandwidth G and threshold k choices directly in captions of key figures. \\
\noindent \textcolor{blue}{\bf Response}: \textcolor{black}{We have added the values of $G$ and $k$ for MOPED to the figure and table captions.}\\
\item Consider moving large simulation tables to an appendix or supplementary material. \noindent \textcolor{blue}{\bf Response}: \textcolor{black}{We have moved Table 2 to the Appendix.}\\
\item In EEG case study figures, add a clear legend or annotation explaining seizure labels and expert annotations.  \\ \textcolor{blue}{\bf Response}: \textcolor{black}{To avoid overcrowding in the figures, we have opted not to include these annotations. }\\
\end{itemize}
}

\bibliographystyle{apalike}
{\small
\bibliography{ref}
}
\end{document}
