---
title: "Conditional extremes analysis of Irish rainfall and windspeed data"
author: "Paddy O'Toole"
date: "2024-07-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, dev = "pdf")
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

## Libs & metadata

```{r, results = 'hide', message = FALSE}
library(sf)
library(dplyr, quietly = TRUE)
library(ggplot2)
library(lubridate)
library(texmex)
library(gridExtra)
library(patchwork)

source("src/functions.R")

theme <- theme_bw() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 16),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 13, face = "bold"),
    strip.background = element_rect(fill = NA, colour = "black"),
    plot.tag = element_text(size = 16, face = "bold"),
    panel.background = element_rect(fill = NA, colour = "black")
  )

min_max_dates <- as_date(c("1990-01-01", "2020-12-31"))
all_dates <- seq.Date(min_max_dates[1], min_max_dates[2], by = "day")
```


## Load Data

Load data on Irish rainfall and windspeeds at various sites, with 
associated shapefiles. 
```{r}
data <- readr::read_csv(
  "data/met_eireann/final/met_eir_wind.csv.gz",
  show_col_types = FALSE
) %>%
  filter(date %in% all_dates, !is.na(wind_speed))

# load shapefile
areas <- sf::read_sf("data/met_eireann/final/irl_shapefile.geojson")
```

## Exploration 

Pairs plots for areas with low (Aughnasheelan (Miskawn)) and higher asymptotic 
extremal dependence (Loughglinn).

```{r}
data %>% 
  filter(name == "Aughnasheelan (Miskawn)") %>% 
  ggplot(aes(x = rain, y = wind_speed)) + 
  geom_point() + 
  geom_hline(yintercept = 10.91, colour = "red") + 
  geom_vline(xintercept = 2, colour = "blue") + 
  theme + 
  ggtitle("Rainfall vs Wind in Aughnasheelan")
```

```{r}
data %>% 
  filter(name == "Loughglinn") %>% 
  ggplot(aes(x = rain, y = wind_speed)) + 
  geom_point() + 
  geom_hline(yintercept = 7.024, colour = "red") + 
  geom_vline(xintercept = 5.3, colour = "blue") + 
  theme + 
  ggtitle("Rainfall vs Wind in Loughglinn")
```

Seem to be very similar? Although a and b are both 0.01 for 
Aughnasheelan (Miskawn), but 0.321 and 0.265 for Loughglinn. 
A lot more extremal observations for Loughlinn though.
windspeeds.

## Fit CE model

Create a matrix of rainfall vs windspeed for each site:
```{r}
mat_list <- data %>% group_split(name) %>% purrr::map(\(x) {
    mat <- as.matrix(select(x, rain, wind_speed))
    colnames(mat) <- c("rain", "windspeed")
    return(mat)
})
names(mat_list) <- unique(data$name)
```

For each site, fit marginal models (i.e. GPD over some threshold):
```{r}
marginal <- lapply(mat_list, migpd, mqu = .95, penalty = "none")
```

For each site, fit dependence models:
```{r}
# fit for each site
dependence <- lapply(marginal, \(x) {
  # fit for rain and wind speed
  lapply(c("rain", "windspeed"), \(col) {
    mexDependence(x, which = col, dqu = 0.95)
  })
})
```

## Dependence

Want to plot a and b for rain and windspeed for each site. 

First, must extract a and b values for each site and preprocess plot data:
```{r}
# pull a and b values for each site
a_b_vals <- lapply(dependence, \(x) {
 list(
   summary(x[[1]])$dependence[1:2],
   summary(x[[2]])$dependence[1:2]
 )
})

# convert to dataframe
a_b_df <- bind_rows(lapply(seq_along(a_b_vals), \(i) {
  data.frame(
    "name" = names(a_b_vals)[i],
    "var"  = c("rain", "rain", "windspeed", "windspeed"),
    "parameter" = c("a", "b", "a", "b"),
    "value" = c(
      a_b_vals[[i]][[1]][1], a_b_vals[[i]][[1]][2],
      a_b_vals[[i]][[2]][1], a_b_vals[[i]][[2]][2]
    )
    # "a"    = c(a_b_vals[[i]][[1]][1], a_b_vals[[i]][[2]][1]),
    # "b"    = c(a_b_vals[[i]][[1]][2], a_b_vals[[i]][[2]][2])
  )
}))

# join in area information, change to sf
a_b_sf <- left_join(
  a_b_df, 
  distinct(data, name, lon, lat)
) %>% 
  st_as_sf(coords = c("lon", "lat"))
st_crs(a_b_sf) <- "WGS84"
```

Now plot:
```{r, fig.width=16, fig.height=12}
p_lst <- lapply(c("a", "b"), \(x) {
  ggplot() +
    geom_sf(data = areas, colour = "black") +
    geom_sf(
      data = filter(a_b_sf, parameter == x),
      # aes(fill = value, size = value),
      aes(fill = value, size = value),
      colour = "black",
      stroke = 1,
      pch = 21
    ) +
    facet_wrap(var ~ parameter, ncol = 2) + 
    scale_size_continuous(
      breaks = seq(0, 0.5, by = 0.1),
      labels = c("0", "0.1", "0.2", "0.3", "0.4", "0.5")
    ) +
    scale_fill_gradientn(
      colours = RColorBrewer::brewer.pal(name = "Blues", n = 7),
      breaks = seq(0, 0.5, by = 0.1),
      labels = c("0", "0.1", "0.2", "0.3", "0.4", "0.5"),
      guide = "legend"
    ) +
    guides(fill = guide_legend(), size = guide_legend()) +
    labs(fill = "", size = "") +
    theme
})

p_lst[[1]] / p_lst[[2]]
```

Positive asymptotic dependence of various strengths is found for rainfall 
and wind speed across all sites. 
Rainfall appears to have greater asymptotic extremal dependence on wind speed
than vice versa for most regions. 

## Marginal

Show summaries for rainfall and wind for two sites previously chosen:
```{r}
summary(marginal$`Aughnasheelan (Miskawn)`)
```

```{r}
summary(marginal$Loughglinn)
```

Both have infinite upper end points for rain, but Loughlinn has much lower 
threshold for windspeeds, giving much more extremal observations. 

Similarly, extract 
First, must extract sigma and xi values for each site and preprocess plot data:
```{r}
# TODO: Functionalise
# pull xi and sigma values for each site
gpd_vals <- lapply(marginal, \(x) {
 list(
   summary(x)$co[3:4, 1],
   summary(x)$co[3:4, 2]
 )
})

# convert to dataframe
gpd_df <- bind_rows(lapply(seq_along(gpd_vals), \(i) {
  data.frame(
    "name" = names(gpd_vals)[i],
    "var"  = c("rain", "rain", "windspeed", "windspeed"),
    "parameter" = c("sigma", "xi", "sigma", "xi"),
    "value" = c(
      gpd_vals[[i]][[1]][1], gpd_vals[[i]][[1]][2],
      gpd_vals[[i]][[2]][1], gpd_vals[[i]][[2]][2]
    )
    # "a"    = c(gpd_vals[[i]][[1]][1], gpd_vals[[i]][[2]][1]),
    # "b"    = c(gpd_vals[[i]][[1]][2], gpd_vals[[i]][[2]][2])
  )
}))

# join in area information, change to sf
gpd_sf <- left_join(
  gpd_df, 
  distinct(data, name, lon, lat)
) %>% 
  st_as_sf(coords = c("lon", "lat"))
st_crs(gpd_sf) <- "WGS84"
```

Now plot sigma values:
```{r, fig.width=16, fig.height=12}
ggplot() +
  geom_sf(data = areas, colour = "black") +
  geom_sf(
    data = filter(gpd_sf, parameter == "sigma"),
    # aes(fill = value, size = value),
    aes(fill = value, size = value),
    colour = "black",
    stroke = 1,
    pch = 21
  ) +
  facet_wrap(~var, ncol = 2) + 
  ggtitle("Sigma values") + 
  scale_size_continuous(
    # breaks = seq(0, 0.5, by = 0.1),
    breaks = seq(0, 12, by = 3),
    # labels = c("0", "0.1", "0.2", "0.3", "0.4", "0.5")
    labels = c("0", "3", "6", "9", "12")
  ) +
  scale_fill_gradientn(
    colours = RColorBrewer::brewer.pal(name = "Blues", n = 7),
    # breaks = seq(0, 0.5, by = 0.1),
    breaks = seq(0, 12, by = 3),
    # labels = c("0", "0.1", "0.2", "0.3", "0.4", "0.5"),
    labels = c("0", "3", "6", "9", "12"),
    guide = "legend"
  ) +
  guides(fill = guide_legend(), size = guide_legend()) +
  labs(fill = "", size = "") +
  theme
```

and xi values:
```{r, fig.width=16, fig.height=12}
# TODO: Make deliniation between positive and negative values more obvious
ggplot() +
  geom_sf(data = areas, colour = "black") +
  geom_sf(
    data = filter(gpd_sf, parameter == "xi"),
    # aes(fill = value, size = value),
    aes(fill = value, size = value),
    colour = "black",
    stroke = 1,
    pch = 21
  ) +
  facet_wrap(~var, ncol = 2) + 
  ggtitle("Xi values") + 
  scale_size_continuous(
    # breaks = seq(0, 0.5, by = 0.1),
    # labels = c("0", "0.1", "0.2", "0.3", "0.4", "0.5")
    breaks = round(seq(-0.3, 0.3, by = 0.1), 1), 
  ) +
  scale_fill_gradient2(
    low = "blue3",
    high = "red3",
    na.value = "grey",
    breaks = round(seq(-0.3, 0.3, by = 0.1), 1) 
    # label = scales::label_percent(accuracy = 1, trim = TRUE)
  ) + 
  guides(fill = guide_legend(), size = guide_legend()) +
  labs(fill = "", size = "") +
  theme
```

Appears that there are some positive shape parameter estimates here for both
rain and windspeed, which would result in our GPD having an infinite upper 
end point, is that realistic?
```{r}
list(
  "rain" = summary(gpd_sf[gpd_sf$parameter == "xi" & gpd_sf$var == "rain", ]$value),
  "ws"   = summary(gpd_sf[gpd_sf$parameter == "xi" & gpd_sf$var == "windspeed", ]$value)
)
```


## Diagnostics

```{r}
ggplot(dependence[[1]])
```

