\begin{document}
% Jordan Richards, Ph.D.
Patrick O'Toole
Department of Mathematical Sciences,
University of Bath

October 22, 2025


Dear Editor,

We are pleased to submit an original research article entitled ``Clustering of multivariate tail dependence using conditional methods'' for consideration for publication in the Journal of Computational and Graphical Statistics.

In this manuscript, we propose a novel framework for clustering multivariate tail dependence structures using conditional extremes models. Our methodology leverages the strengths of the widely-used conditional extremes model to provide a flexible yet parsimonious representation of multivariate tail dependence.
We cluster the parameters of the conditional extremes model using the skew-geometric Jensen-Shannon divergence, which provides a closed-form, compututionally efficient measure of dissimilarity between two models. This allows us to effectively capture and group similar tail dependence structures in high-dimensional settings.
We demonstrate the efficacy of our approach through an extensive simulation study, showing that it outperforms existing methods in accurately recovering underlying tail dependence structures, and uniquely extends to $d>2$ dimensions.
We further illustrate the practical utility of our methodology by applying it to a real-world dataset involving precipitation and wind speed extremes in the Republic of Ireland, revealing insightful patterns in the joint extremal behaviour of these environmental variables.

Given our novel methodology, extensive simulation study, and effective real-world application, we believe that the manuscript will be of interest to JCGS readers.

% Alongside the substantial methodological advancements, we have provided an early version of an R package that we believe has the potential to become the industry standard for fitting regression models. This manuscript has not been published and is not under consideration for publication elsewhere. We have no conflicts of interest to disclose. If the paper is thought to be too long, we are also happy to move part of the application to the supplementary material.
An R package implementing the proposed methodology is in preparation and will soon be available for reviewers and readers. 
This manuscript has not been published and is not under consideration for publication elsewhere. 
We have no conflicts of interest to disclose. 

Thank you for your consideration! We look forward to hearing from you.

Sincerely,

Patrick O'Toole, Christian Rohrbeck, and Jordan Richards

\end{document}
